---
title: "Identify interesting genes in different analyses"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(AnnotationDbi)
```

## The purpose of this script is to see which of the genes in modules, DGE, and GSEA analyses could be candidates for the effect of diet (i.e. diet-tissue interaction)

# Are known pathway genes present in modules?

# Map modules to table of gene summaries

```{r}

# Genes from QTL
qtgs <- read_csv(file = "../../processed/DESEQ/DEG_QTL/unique_nearGenes.csv") %>%
  as.data.frame() %>%
  dplyr::select(2)

colnames(qtgs)[colnames(qtgs)=="x"] <- "gene_symbol"




# Genes from WGCNA

# Load module names
load("../../processed/DESEQ/Coexpression/modules.RData")

# Read in multiple files by wildcard expansion on file paths
getMods <- lapply(Sys.glob(
  "../../processed/DESEQ/Coexpression/Mods/gene_symbols/*.txt"),read.table)

# Rename column
ss <- "gene_symbol"
for (i in 1:length(getMods)) {
 colnames(getMods[[i]]) <- ss 
}

# Add column for module name to each module
getMods <- mapply(cbind, getMods, "mod"=sort(modNames), 
                  SIMPLIFY=FALSE)
 
# Merge modules in list into one table
ll <- bind_rows(getMods)
ll$FBgn = mapIds(org.Dm.eg.db,
                     keys=as.character(ll$gene_symbol), 
                     column="FLYBASE",
                     keytype="SYMBOL",
                     multiVals="first")



sigDEGs <- read.csv("../../processed/DESEQ/DEGs_lrt.treatment_0.05.csv")
head(sigDEGs)

sigDEGs$symbol = mapIds(org.Dm.eg.db,
                     keys=as.character(sigDEGs$Gene), 
                     column="SYMBOL",
                     keytype="FLYBASE",
                     multiVals="first")

# Pathway and GO term genes
pgo <- read.csv("../../processed/DESEQ/gopath_genes.csv")                     
```

# Intersection between QTL and module analysis

```{r}
# Common between QTL interval genes and DEGs for treatment
common <- as.data.frame(intersect(qtgs$gene_symbol, ll$gene_symbol))
colnames(common) <- "symbol"

common$FBgn = mapIds(org.Dm.eg.db,
                     keys=as.character(common$symbol), 
                     column="FLYBASE",
                     keytype="SYMBOL",
                     multiVals="first")

# In which modules are these?
common_mod <- subset(ll,ll$gene_symbol %in% common)


# Look at FC and significance
comm_mod_exprs <- subset(sigDEGs,sigDEGs$symbol %in% common)

# Associated GO and pathway terms
comm_mod_gopath <- subset(gopath_genes,gopath_genes$symbol %in% common)

xx <- read.csv("../../processed/DESEQ/Coexpression/minMod30/GOEnrichmentTable.csv")
 colnames(xx)[colnames(xx)=="module"] <- "mod"

# Combine into table
colnames(common_mod)[colnames(common_mod)=="gene_symbol"] <- "symbol"
joint <- inner_join(common_mod,comm_mod_exprs, by="symbol")
joint <- inner_join(joint, comm_mod_gopath, by="symbol")
# joint <- inner_join(joint, xx, by="mod")

# This list has all genes under QTL - not limiting to 100 Mb
qtall <- read.csv("../../processed/DESEQ/DEG_QTL/all_diffExpr_underQTL.csv")
str(qtall)

qtmod <- subset(ll,ll$gene_symbol %in% qtall$gname)
colnames(qtmod)[colnames(qtmod)=="gene_symbol"] <- "symbol"
qtmod_gopath <- subset(gopath_genes,gopath_genes$symbol %in% qtall$gname)

joint1 <- inner_join(qtmod, qtmod_gopath, by="symbol")
joint2 <- inner_join(joint1, qtmod_sigDEGs, by="symbol")

# Common between QTL interval genes and DEGs for treatment
common1 <- intersect(sigDEGs$symbol, qtgs$gene_symbol)
```

# Intersection between QTL and DEGs with FC

```{r}
qtmod_sigDEGs <- subset(sigDEGs,sigDEGs$symbol %in% qtall$gname) 
                          
# Map to fold change table
qgs <- dplyr::select(qtmod_sigDEGs, Gene,symbol)
fc <- as.data.frame(all.fc.dat) %>%
  rownames_to_column(var="Gene") %>%
    dplyr::select(-c(contains("FCse_"),
                   contains("stat_"),
                   contains("p_"),
                   contains("padj_")),-"log2FoldChange")

qgs <- inner_join(qgs,fc)
qgss <- subset(ll,ll$gene_symbol %in% qgs$symbol)
colnames(qgss)[colnames(qgss)=="gene_symbol"] <- "symbol"
qgs <- inner_join(qgs,qgss)
qgs <- qgs %>% mutate_if(is.numeric, round, digits = 6)



```

# Sirtuins and FOXO pathway genes

```{r}
# Ref: https://flybase.org/lists/FBgg/pathways

# FOXO signaling pathway (i.e. including sirtuins)
foxo_sirts <- read.csv("../../processed/DESEQ/DEG_QTL/FOXO_sigPath.csv")
fox <- foxo_sirts[,c(1,3,4)]
#colnames(foxo_sirts)[1]<-"FBgn"
fox <- subset(ll,ll$FBgn %in% foxo_sirts$primary_FBgn)
fox1 <- subset(foxo_sirts,foxo_sirts$primary_FBgn %in% common$FBgn)
fox2 <- subset(sigDEGs, sigDEGs$Gene %in% foxo_sirts$primary_FBgn)
fox2$pathway <- "foxoSirt"
```

1. Out of 105 FOXO genes plus 5 sirtuins, modules hold 109 of these (fox)
2. Zero QTL genes and zero sigDEgs map to this list (fox1)
3. 16 FOXO and sirtuins were DE (fox2: Sirt2, Diap1, beta'COP, CG5059, Gycbeta1008, DOR, TBPH, hid,p38c, Ilp5, alphaCOP, AMPKalpha, Rheb, Cdk2, arm, Atg1)


```{r}
# TOR pathway
tor <- read.csv("../../processed/DESEQ/DEG_QTL/TOR_path.csv")
tor <- tor[,c(1,3,4)]

tor1 <- subset(ll,ll$FBgn %in% tor$primary_FBgn) # list not screened further
tor2 <- subset(tor,tor$primary_FBgn %in% common$FBgn)
tor3 <- subset(sigDEGs, sigDEGs$Gene %in% tor$primary_FBgn)
tor3$pathway <- "TOR"

```

1. 151 of 152 TOR genes found in modules
2. Zero of QTL genes
3. 25 DEGs are in TOR pathway: Vha68-1, tsl, Socs36E, Hsc70-4, sev, CG16908, tub, dpp, eIF4E1, hid, svp, unk, Wdr24, AMPKalpha, stumps, htl, gcl, Rheb, Akap200, arm, slmb, Myc, fkh, Nup44A, Atg1 

```{r}
# Toll pathway genes
toll <- read.csv("../../processed/DESEQ/DEG_QTL/tollPath_precomp.csv")
toll <- toll[,c(1,3,4)]

toll1 <- subset(ll,ll$FBgn %in% toll$primary_FBgn) # list not screened further
toll2 <- subset(toll,toll$primary_FBgn %in% common$FBgn)
toll3 <- subset(sigDEGs, sigDEGs$Gene %in% toll$primary_FBgn)
toll3$pathway <- "Toll"
```

1. 49 Toll genes recovered in modules. List not checked further (toll1)
2. Zero of 12 QTL genes found (toll2)
3. 12 DEGs are Toll pathway genes (toll3: PGRP-SC1a, spz, Gprk2, tub, dgrn, grass, wek, Spn88Ea, senju, spirit, Hrs, Herc4)

```{r}
# IIS pathway genes
iis <- read.csv("../../processed/DESEQ/DEG_QTL/iisPath_precomp.csv")
iis <- iis[,c(1,3,4)]

iis1 <- subset(ll,ll$gene_symbol %in% iis$annotation_ID)
iis2 <- subset(iis,iis$primary_FBgn %in% common$FBgn)
iis3 <- subset(sigDEGs, sigDEGs$Gene %in% iis$primary_FBgn)
iis3$pathway <- "IIS"
```

1. 55 IIS pathway genes found in modules (iis1)
2. Zero of 12 QTL genes are IIS (iis2)
3. 6 DEGs are IIS pathway genes (iis3: Ilp5, tgo, Rheb, slmb, Myc, lin-28)

```{r}
indy <- read.csv("../../processed/DESEQ/DEG_QTL/indy.csv")
indy <- indy[c(1,3,4)]

indy1 <- subset(ll,ll$gene_symbol %in% indy$annotation_ID)
indy2 <- subset(common, common$FBgn %in% indy$primary_FBgn)
indy3 <- subset(sigDEGs, sigDEGs$Gene %in% indy$primary_FBgn)
indy1 <- subset(ll, ll$Gene=="FBgn0052027")
indy3$pathway <- "Indy"
```

# Candidates

```{r}
candids <- rbind(fox2,tor3,iis3)
sigs <- sigDEGs[,c(1,8)]
lll <-ll%>%
  rename(symbol=gene_symbol,
         Gene=FBgn)

candidates <- inner_join(candids,sigs)
candidates <- inner_join(candidates,lll)
candidates$Gene <- as.character(candidates$Gene)

# Merge with shrinked FC data
cgs <- dplyr::select(candidates, Gene,symbol,pathway,mod)
fc <- as.data.frame(all.fc.dat) %>%
  rownames_to_column(var="Gene")
cgs <- inner_join(cgs,fc) %>%
  dplyr::select(-c(contains("FCse_"),
                   contains("stat_"),
                   contains("p_"),
                   contains("padj_")),-"log2FoldChange")

cgs <- cgs %>% mutate_if(is.numeric, round, digits = 6)

write.csv(cgs, file = "../../processed/DESEQ/cand_genes.csv" )
```

47 hits in DEGs (i.e. 38 unique genes)

# Gene summaries
```{r}
gsum <- read_lines(file = "../../../../automated_gene_summaries.tsv",
                   skip = 1)%>%
  str_split(pattern = "\t", simplify = TRUE) %>% 
  as_tibble() 
gsum <- gsum[,c(1,2)]
colnames(gsum) <- c('FBgn', 'summary')

candsum <- subset(gsum, gsum$FBgn %in% candidates$FBgn)
#colnames(candidates)[colnames(candidates)=="Gene"] <- "FBgn"
candsum <- inner_join(candsum,candidates, by="FBgn") %>%
  dplyr::select(c(1,9,everything())) %>% as.data.frame()

```

