---
title: "Module eigengene expression across diets"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(cowplot)

```

# Plot gene expression for each module
This script subsets and plots gene expression for each WGCNA module. Use data corrected for batches with removeBatchEffects() i.e. "../../processed/DESEQ/rlogtrt.csv" created in wgcnaCoex.Rmd


# Expression data

```{r}
# Eigengenes from wgcnaCoex.Rmd
temp_wide <- read.csv("../../processed/DESEQ/rlogtrt_batchCor.csv")
names(temp_wide)[1] <- "FBgn"           

# Reformat to long form
temp.long <-gather(temp_wide, Diet, Expression, C1_B:HS6_O, factor_key=FALSE)
                    
# Split Diet column
temp.long <- separate(temp.long, 
                      col = Diet, 
                      into = c("Diet", "Tissue"), 
                      sep = "_") 

temp.long$FBgn <- as.character(temp.long$FBgn)

```

# Get module genes

```{r}
# Modules & module names from WGCNA
load("../../processed/DESEQ/Coexpression/modules.RData")

# Get module genes
modNames
table(moduleColors)

getMods <- modNames[1:22]

for (module in getMods) {
  # select module genes
  modGenes = (moduleColors==module)
  # Get their FBgn ids
  modIDs = temp.long$FBgn[modGenes]
  # Write file
  fileName = paste("../../processed/DESEQ/Coexpression/Mods/", 
                   module,".txt", sep="")
  write.table(as.data.frame(modIDs), file = fileName,
              row.names = FALSE, col.names = FALSE)
}

```

# Subset expression data using module genes
# For one module

```{r}
# Read module
cyan <- read.table("../../processed/DESEQ/Coexpression/Mods/cyan.txt",
                   header = FALSE, stringsAsFactors = FALSE) 
names(cyan)[1] <- "FBgn"

cyan <- inner_join(cyan, temp.long, by="FBgn")

cy_mean <- cyan %>%
  group_by(Diet, Tissue) %>%
  summarize(mean_expr = mean(Expression, na.rm=TRUE))
cy_mean$Diet1 <- gsub('[0-9]+', '', cy_mean$Diet)

ggplot(cy_mean,aes(Diet1, mean_expr, group=1, color=Tissue)) +
  geom_point(position = position_jitter(width = 0.15)) +
  geom_point(stat = "summary", fun.y=mean) +
  stat_summary(fun.y=mean, geom ="line")

```

# Put modules gene lists in one table

```{r}
getMods <- lapply(Sys.glob(
  "../../processed/DESEQ/Coexpression/Mods/*.txt"),read.table)

# Rename column
ss <- "FBgn"
for (i in 1:length(getMods)) {
 colnames(getMods[[i]]) <- ss 
}


# Add column for module name to each module
getMods <- mapply(cbind, getMods, "mod"=sort(modNames), 
                  SIMPLIFY=FALSE)
 
# Merge modules in list into one table
ll <- bind_rows(getMods)


# Merge module lists with expression data
lk <- left_join(temp.long,ll, by="FBgn")

# Test on one module
mm <- subset(lk, lk$mod=="cyan")
i.mm_mean <- mm %>%
  group_by(Diet, Tissue) %>%
  summarize(mean_expr = mean(Expression, na.rm=TRUE))
i.mm_mean$Diet1 <- gsub('[0-9]+', '', i.mm_mean$Diet)

i.mm_plot <- ggplot(i.mm_mean,aes(Diet1, mean_expr, group=1, color=Tissue)) +
  geom_point(position = position_jitter(width = 0.15)) +
  geom_point(stat = "summary", fun.y=mean) +
  stat_summary(fun.y=mean, geom ="line")
i.mm_plot
```

# Summarize and plot

```{r}
# New grouping column
lk$Diet1 <- gsub('[0-9]+', '', lk$Diet)

lk_mean <- lk %>%
  group_by(mod,Tissue,Diet1) %>%
  summarize(mean_expr = mean(Expression, na.rm=TRUE))

yf <- unique(lk_mean$mod)

mod.plots <- vector(mode="list", length=length(yf))
names(mod.plots) <- yf

for(yy in yf)
{
  
ff <- ggplot(lk_mean,aes(Diet1, mean_expr, group=1, color=Tissue)) +
  #geom_point(position = position_jitter(width = 0.15)) +
  geom_point(stat = "summary", fun.y=mean) +
  stat_summary(fun.y=mean, geom ="line") +
  #stat_summary(fun.data = mean_se, geom = "errorbar") +
  theme(legend.position="none") +
  labs(x = "Diet", y = "Mean expression")

ff 
  
mod.plots[[yy]] <- ff

}

#mod.plots[["darkgreen"]]

modplot <- plot_grid(plotlist=mod.plots,nrow=5,rel_widths = c(1.3,1.1,1.8),rel_heights = c(1,1,1.2), label_size = 9) 

modplot
#legend <- get_legend(modplot)

modplot1 <- plot_grid(modplot, legend, rel_widths = c(3,3))
```

