---
title: "Confirmatory GO analysis via DAVID"
author: "Enoch Ng'oma"
date: "10/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(AnnotationDbi)
library(org.Dm.eg.db)
```

# Data

```{r}
# Background List: 12614 genes from DESeq

load("../../processed/DESEQ/all_fc_dat.rda")
genes2go <- as.data.frame(all.fc.dat)

# List to test enrichment: 2475 genes
sigDEGs <- read.csv("../../processed/DESEQ/DEGs_lrt.treatment_0.05.csv",
                    stringsAsFactors = FALSE)
sigDEGs$Gene <- gsub('FBgn', 'FBGN', sigDEGs$Gene)

# Get both lists to DAVID online v.6.8
genes2go <- genes2go$X
sigs <- sigDEGs$Gene
```
Ref: https://www.biostars.org/p/299161/
Procedure online
1. Upload list of 2475 genes to enrich
2. Upload background list of 12614 - all genes with >1 transcript (/DESEQ/rlogtrt_batchCor.csv)
3. Select the identifier - FLYBASE
4. Select 'Gene List'
5. Select 'Bakground List'
6. Select all repositories against to perform enrichment (i.e. GO BP, MF, and CC) 
7. Select Functional Annotation Chart


# Enrichment analysis

```{r}
FC <- read.csv("../../processed/DESEQ/GO/David_all_lrtTreat_fbgn.csv", 
                 stringsAsFactors = FALSE)

sigGeneList <- FC

#log2FCcutoff <- 1.5 #GO_1
log2FCcutoff <- 1.0 #GO_2
BHcutoff <- 0.05
sigGeneList <- subset(sigDEGs, abs(log2FoldChange)>=log2FCcutoff & padj<=BHcutoff)[,1]

# Enrichment cut-off
enrichBcutoff <- 0.05
DAVID <- subset(FC, Benjamini<enrichBcutoff)
DAVID <- DAVID[,c(1,2,3,6)]

# Create a new dataframe that has '1' for when the gene is part of a term, and '0' when not
annGSEA <- data.frame(row.names=sigGeneList)
for (j in 1:length(sigGeneList))
{
    gene <- sigGeneList[j]
    pattern <- paste("^", gene, ", |, ", gene, "$| ", gene, ",", sep="")
    for (k in 1:nrow(DAVID))
    {
        if (any(grepl(pattern, DAVID$Genes[k])))
        {
            annGSEA[j,k] <- 1
        }
        else
        {
            annGSEA[j,k] <- 0
        }
    }
}
colnames(annGSEA) <- DAVID[,2]
  
  #Remove terms with no overlapping genes
  annGSEA <- annGSEA[,apply(annGSEA, 2, mean)!=0]
  #Remove genes with no overlapping terms
  annGSEA <- annGSEA[apply(annGSEA, 1, mean)!=0,]
```

# Generate heatmap and annotation

```{r}
require(ComplexHeatmap)
require(circlize)

#Match the order of rownames in toptable with that of annGSEA
sigDEGs <- sigDEGs[which(sigDEGs$Gene %in% rownames(annGSEA)),]
sigDEGs <- sigDEGs[match(rownames(annGSEA), sigDEGs$Gene),]

#Set text and figure dimensions
geneLab=10
termLab=8

# Conver ids in both objects
rownames(annGSEA) <- gsub('FBGN', 'FBgn', rownames(annGSEA))
rownames(annGSEA) = mapIds(org.Dm.eg.db,
                     keys=as.character(rownames(annGSEA)), 
                     column="SYMBOL",
                     keytype="FLYBASE",
                     multiVals="first")

sigDEGs$Gene <- gsub('FBGN', 'FBgn', sigDEGs$Gene)
sigDEGs$Gene = mapIds(org.Dm.eg.db,
                     keys=as.character(sigDEGs$Gene), 
                     column="SYMBOL",
                     keytype="FLYBASE",
                     multiVals="first")


#Create heatmap annotations
    #Colour bar for -log (base 10) FDR Q value for DEGs, and fold changes
    dfMinusLog10FDRGenes <- data.frame(-log10(sigDEGs[which(sigDEGs[,1] %in% rownames(annGSEA)),"padj"]))
    dfMinusLog10FDRGenes[dfMinusLog10FDRGenes=="Inf"] <- 0
    dfFoldChangeGenes <- data.frame(sigDEGs[which(sigDEGs[,1] %in% rownames(annGSEA)),"log2FoldChange"])
    dfGeneAnno <- data.frame(dfMinusLog10FDRGenes, dfFoldChangeGenes)
    colnames(dfGeneAnno) <- c("DEG\nsignificance\nscore", "Regulation")
    dfGeneAnno[,2] <- ifelse(dfGeneAnno[,2]>0, "Up-regulated", "Down-regulated")
    colours <- list("Regulation"=c("Up-regulated"="royalblue", "Down-regulated"="yellow"))
    haGenes <- rowAnnotation(df=dfGeneAnno, col=colours, width=unit(1,"cm"))

    #Colour bar for -log (base 10) Benjamini enrichment Q value
    dfMinusLog10BenjaminiTerms <- data.frame(-log10(read.csv("../../processed/DESEQ/GO/David_all_lrtTreat_fbgn.csv", header=TRUE)[which(read.csv("../../processed/DESEQ/GO/David_all_lrtTreat_fbgn.csv", header=TRUE)$Term %in% colnames(annGSEA)),"Benjamini"]))
    colnames(dfMinusLog10BenjaminiTerms) <- "GO Term\nsignificance\nscore"
    haTerms <- HeatmapAnnotation(df=dfMinusLog10BenjaminiTerms,
        colname=anno_text(colnames(annGSEA), rot=40, just="right", offset=unit(1,"npc")-unit(2,"mm"), gp=gpar(fontsize=termLab)),
        annotation_height=unit.c(unit(1, "cm"), unit(8, "cm")))

pdf("../../plots/DAVID_GO_2.pdf", width=7, height=6)
    hmapGSEA <- Heatmap(annGSEA,

        name="My enrichment",

        split=dfGeneAnno[,2],

        col=c("0"="white", "1"="forestgreen"),

        rect_gp=gpar(col="grey85"),

        cluster_rows=T,
        show_row_dend=T,
        row_title="Statistically-significant genes",
        row_title_side="left",
        row_title_gp=gpar(fontsize=12, fontface="bold"),
        row_title_rot=0,
        show_row_names=TRUE,
        row_names_gp=gpar(fontsize=geneLab, fontface="bold"),
        row_names_side="left",
        row_names_max_width=unit(15, "cm"),
        row_dend_width=unit(10,"mm"),

        cluster_columns=T,
        show_column_dend=T,
        column_title="Enriched terms",
        column_title_side="top",
        column_title_gp=gpar(fontsize=12, fontface="bold"),
        column_title_rot=0,
        show_column_names=FALSE,
        #column_names_gp=gpar(fontsize=termLab, fontface="bold"),
        #column_names_max_height=unit(15, "cm"),

        show_heatmap_legend=FALSE,

        #width=unit(12.5, "cm"),

        clustering_distance_columns="euclidean",
        clustering_method_columns="ward.D2",
        clustering_distance_rows="euclidean",
        clustering_method_rows="ward.D2",

        bottom_annotation=haTerms)

    draw(hmapGSEA + haGenes, heatmap_legend_side="right", annotation_legend_side="right")
dev.off()
```

