---
title: "Gene set enrichment analysis (GSEA)"
author: "Enoch Ng'oma"
date: "12/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(DESeq2)
library(tidyverse)
library(cowplot)
library(gage)
library(pathview)
library(AnnotationDbi)

# Get Dmel gene annotation database
library(org.Dm.eg.db)
library(GO.db)
```

# Set up kegg and GO gene sets

```{r}
columns(org.Dm.eg.db)

# Generate up-to-date KEGG pathway gene sets 
kegg_sets <- kegg.gsets(species = "dme", id.type = "entrez", check.new=FALSE)

# check that expr data and gene sets both use entrez
lapply(kegg_sets[1:3],head)
head(kegg_sets[[1]])

# sigmet.idx: signaling or metabolisms pathways
kegg_sets.dm = kegg_sets$kg.sets[kegg_sets$sigmet.idx]
head(kegg_sets.dm, 3)

#kegg.sets.dm = id.map.fb[kegg.sets$sigmet.idx]

# Generate up-to-date GO (Gene Ontology) gene sets
go_sets <- go.gsets(species = "Fly", 
                    pkg.name="org.Dm.eg.db", 
                    id.type = "eg", 
                    keep.evidence=FALSE)

# Do GO on Biological Process only (i.e. see ...$go.subs)
gobpsets = go_sets$go.sets[go_sets$go.subs$BP]
```

# List of DEGs for treatment
- GSEA accounts for expression data to calculate enrichment.

```{r}
# List of genes significant for treatment
lrt.treatment <- read.csv(file="../../processed/DESEQ/DEGs_lrt.treatment_0.05.csv")
# rownames(lrt.treatment) <- lrt.treatment[,1]
# lrt.treatment[,1] <- NULL
lrt.treatment <- lrt.treatment[,1]
lrt.treatment[1:4]

load(file="../../processed/DESEQ/all_fc_dat.rda")

lrt.treat <- subset(all.fc.dat,rownames(all.fc.dat) %in% lrt.treatment)
```

# Need libraries: gage, pathview and org.Dm.eg.db
# Used this tutorial together with all 3 vignettes: http://www.gettinggeneticsdone.com/2015/12/tutorial-rna-seq-differential.html

```{r}
# Add annotation maps to results object

# gene symbol
lrt.treat$symbol = mapIds(org.Dm.eg.db,
                     keys=row.names(lrt.treat), 
                     column="SYMBOL",
                     keytype="FLYBASE",
                     multiVals="first")

# entrez id
lrt.treat$entrez = mapIds(org.Dm.eg.db,
                     keys=row.names(lrt.treat), 
                     column="ENTREZID",
                     keytype="FLYBASE",
                     multiVals="first")

# full gene name
lrt.treat$name =   mapIds(org.Dm.eg.db,
                     keys=row.names(lrt.treat), 
                     column="GENENAME",
                     keytype="FLYBASE",
                     multiVals="first")

#head(lrt.treat)
head(lrt.treat$symbol)

```

# Pathway anaylsis with GAGE
# Aim: enriched pathways

```{r}
cc.sets <- c("FC_DR_B","FC_DR_O","FC_DR_H",
             "FC_HS_B","FC_HS_O","FC_HS_H",
             "FC_DRHS_B","FC_DRHS_O","FC_DRHS_H")

e.names <- lrt.treat$entrez

lrt.treat <- lrt.treat[,cc.sets]

out.list <- vector(mode='list', length=length(cc.sets))
names(out.list)  <- cc.sets

for (ii in 1:length(cc.sets)) 
  {

  fc.i <- lrt.treat[,ii]
  names(fc.i) <- e.names

  fc.kegg = gage(fc.i, gsets = kegg_sets.dm, 
                 use.stouffer = TRUE, same.dir = TRUE) 
  #kegg.sig<-sigGeneSet(fc.kegg)
  # Do both TRUE and FALSE for same.dir
  
out.list[[ii]] <- fc.kegg
}

lapply(out.list[[ii]], head)

# sigGeneSet(out.list[[ii]], cutoff = 0.1,heatmap = FALSE,
#            qpval = c("q.val","p.val")[1],
#            outname="array", pdf.size = c(7,7),
#            p.limit=c(0.5, 5.5), stat.limit=5)

# Note: setting "same.dir=FALSE" yields nothing, and is not recommended (?gage)
```

# Get pathways

```{r}

enrich.path <- data.frame("pathway"=character(length=0),
                          "stat.mean"=numeric(length=0),
                          "q.val"=numeric(length = 0),
                          "contrast"=character(length=0), stringsAsFactors = FALSE)

for(fc.set in cc.sets)
{
  sub.fc <- out.list[[fc.set]][[1]]%>%as.data.frame()%>%add_column(contrast = fc.set)
  
sub.fc <- rownames_to_column(sub.fc, "pathway") %>%
  dplyr::select(pathway,stat.mean,q.val,contrast)

enrich.path <- rbind(enrich.path, sub.fc)
  
}

write.csv(fc.kegg, "../../processed/DESEQ/GO/pathway_enrich.csv")

write.csv(rbind(fc.kegg$greater, fc.kegg$less),
     "../../processed/DESEQ/GO/pathway_enrich.csv")

# outlist.sig <-read.csv("../../processed/DESEQ/pathway_enrich.csv")
# outlist.sig<-sigGeneSet(fc.kegg, outname="outlist.kegg")
```

# Significant pathways

```{r}
# Pathways that differ by diet globally
pqval05 <- subset(enrich.path,enrich.path$q.val<0.05)
pqval05 <- pqval05[order(pqval05$stat.mean, decreasing = TRUE),]
unq05 <- unique(pqval05$pathway)

pqval05$pathway <- c("dme01100: MPs","dme01200: CM","dme00190: OP","dme04141: PP_ER")
xlab <- c("dme04141: PP_ER","dme00190: OP","dme01200: CM","dme01100: MPs")
ggplot(pqval05, aes(x=pathway, y=stat.mean, fill=q.val, label = contrast)) +
  geom_bar(stat = "identity", width = 0.8) + ylab("Geneset level change") +
  xlab("Enriched pathways") + 
  # geom_text(
  #   aes(label = contrast),
  #   position = position_dodge((width = 1)),
  #   vjust = -0.5, hjust = -0.5, size = 3) +
  scale_x_discrete(limits = xlab) +
  theme(axis.text = element_text(size=12)) +
  coord_flip()

pqval1 <- subset(enrich.path,enrich.path$q.val<0.1)

```

# Get genes for enriched pathways

```{r}
paths <- kegg_sets[[1]]
paths1 <- paths[c("dme04141 Protein processing in endoplasmic reticulum", "dme01200 Carbon metabolism","dme01100 Metabolic pathways","dme00190 Oxidative phosphorylation")]

dme_protpER <- as.data.frame(paths1[[1]])
dme_protpER$go_path <- "dme04141"
names(dme_protpER)[1]<-"entrZ"

dme_carbMet <- as.data.frame(paths1[[2]])
dme_carbMet$go_path <- "dme01200"
names(dme_carbMet)[1]<-"entrZ"

dme_metPath <- as.data.frame(paths1[[3]])
dme_metPath$go_path <- "dme01100"
names(dme_metPath)[1]<-"entrZ"

dme_oxiPhos <- as.data.frame(paths1[[4]])
dme_oxiPhos$go_path <- "dme00190"
names(dme_oxiPhos)[1]<-"entrZ"

path_genes <- rbind(dme_protpER,dme_carbMet,dme_metPath,dme_oxiPhos)
path_genes$entrZ <- as.character(path_genes$entrZ)

# Map back to FBgn and symbol
# FBgn id
path_genes$FBgn = mapIds(org.Dm.eg.db,
                     keys=path_genes$entrZ, 
                     column="FLYBASE",
                     keytype="ENTREZID",
                     multiVals="first")
# Gene symbols
path_genes$symbol = mapIds(org.Dm.eg.db,
                     keys=path_genes$entrZ, 
                     column="SYMBOL",
                     keytype="ENTREZID",
                     multiVals="first")

```

# Correlate pathways for pair-contrasts

```{r}
p.statmean <- enrich.path %>% spread(contrast,stat.mean)
head(p.statmean, 10)

subset(p.statmean, p.statmean$q.val<0.05)

# library(GGally)
# 
# ggpairs(p.statmean[,2:9]) 

```

# Gene Ontology (GO)

```{r}
go.list <- vector(mode='list', length=length(cc.sets))

names(go.list)  <- cc.sets

for (kk in 1:length(cc.sets)) 
  {

  fc.k <- lrt.treat[,kk]
  names(fc.k) <- e.names

  gobpres = gage(fc.k, gsets=gobpsets, same.dir=TRUE)
go.list[[kk]] <- gobpres
}

lapply(go.list[[kk]], head)


```

# Get enriched GO terms

```{r}
enrich.goterm <- data.frame("goterm"=character(length=0),
                          "stat.mean"=numeric(length=0),
                          "q.val"=numeric(length = 0),
                          "contrast"=character(length=0), stringsAsFactors = FALSE)

for(fc.set in cc.sets)
{
  sub.fc <- go.list[[fc.set]][[1]]%>%as.data.frame()%>%add_column(contrast = fc.set)
  
sub.fc <- rownames_to_column(sub.fc, "goterm") %>%
  dplyr::select(goterm,stat.mean,q.val,contrast)

enrich.goterm <- rbind(enrich.goterm, sub.fc)
}

go.smean <- enrich.goterm %>% spread(contrast,stat.mean)
head(go.smean, 10)

subset(go.smean,c(go.smean$FC_DR_B & go.smean$goterm ==
                     "GO:0017144 drug metabolic process"))
#library(GGally)
#ggpairs(go.smean[,2:9]) 
```

# Filter GO terms

```{r}
# GO terms that differ by diet globally
its <- strsplit(enrich.goterm$contrast, split = "_", fixed = TRUE)
its <- unlist(lapply(its, function(x) x[1]))
enrich.goterm$diet <- its
enrich.goterm$diet <- gsub('DRHS', 'DH', enrich.goterm$diet)
(enrich.goterm <- enrich.goterm[order(enrich.goterm$q.val), ])
#dup <- enrich.goterm[duplicated(enrich.goterm), ]

# qval < 0.05
gqval05 <- subset(enrich.goterm,enrich.goterm$q.val<0.05) 
gqval05 <- gqval05[order(gqval05$contrast),]
ung05 <- unique(gqval05$goterm)

# qval < 0.01
gqval01 <- subset(enrich.goterm,enrich.goterm$q.val<0.01) 
gqval01 <- gqval01[order(gqval01$contrast),]
ung01 <- unique(gqval01$goterm)

# Find ancestral terms (GO.db)
terms <- c("GO:0044281","GO:0007154","GO:0023052","GO:0007165")
as.list(GOBPANCESTOR[terms]) 

# Note: GOMFANCESTOR for MF, GOCCANCESTOR, for CC, GOCCCHILDREN, for children

# Write table of pathways and GO terms together
gqval05 <- gqval05[,1:4]
colnames(gqval05)[colnames(gqval05)=="goterm"] <- "term"
colnames(pqval1)[colnames(pqval1)=="pathway"] <- "term"
path_terms <- rbind(pqval1,gqval05)
write.csv(path_terms, "../../processed/DESEQ/GO/sig_terms.csv") 

# For parent term analysis in QUICKGO
term_names <- gqval05$term
term_names <- substr(term_names,1, 10)
term_names <- paste(term_names,collapse = ",")
write.table(term_names, file = "../../processed/DESEQ/GO/term_names.txt", row.names = FALSE)
```

# Get term genes

```{r}
gots <- go_sets[[1]]
gots1 <- gots[c("GO:0044281 small molecule metabolic process", "GO:0007154 cell communication","GO:0023052 signaling","GO:0007165 signal transduction")]

go_smMP <- as.data.frame(gots1[[1]])
go_smMP$go_path <- "GO_0044281"
names(go_smMP)[1]<-"entrZ"

go_cellCom <- as.data.frame(gots1[[2]])
go_cellCom$go_path <- "GO_0007154"
names(go_cellCom)[1]<-"entrZ"

signalL <- as.data.frame(gots1[[3]])
signalL$go_path <- "GO_0023052"
names(signalL)[1]<-"entrZ"

go_sigTransd <- as.data.frame(gots1[[4]])
go_sigTransd$go_path <- "GO_0007165"
names(go_sigTransd)[1]<-"entrZ"

go_genes <- rbind(go_smMP,go_cellCom,signalL,go_sigTransd)
go_genes$entrZ <- as.character(go_genes$entrZ)

# Map back to FBgn and symbol
# FBgn id
go_genes$FBgn = mapIds(org.Dm.eg.db,
                     keys=go_genes$entrZ, 
                     column="FLYBASE",
                     keytype="ENTREZID",
                     multiVals="first")
# Gene symbols
go_genes$symbol = mapIds(org.Dm.eg.db,
                     keys=go_genes$entrZ, 
                     column="SYMBOL",
                     keytype="ENTREZID",
                     multiVals="first")

gopath_genes <- rbind(path_genes,go_genes)
unq_gopath <- unique(gopath_genes$FBgn)
write.csv(gopath_genes, file = "../../processed/DESEQ/gopath_genes.csv")
```

