---
title: "DiffExpr_batchRemove"
author: "Enoch Ng'oma"
date: "5/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load R packagees:
library(tidyverse)
library(cowplot)
library(reshape)
#library(ggthemes)
library(sva)
library(pamr)
library(limma)

set.seed(3537633)
```

```{r}
# Function to extract batch-corrected expression values from a sva object

# Ref:
# https://support.bioconductor.org/p/87508/
# https://support.bioconductor.org/p/47350/

cleaningY = function(y, mod, svaobj) {
  X = cbind(mod, svaobj$sv)
  Hat = solve(t(X)%*%X)%*%t(X)
  beta = (Hat%*%t(y))
  P = ncol(mod)
  cleany = y-t(as.matrix(X[,-c(1:P)])%*%beta[-c(1:P),]) 
  return(cleany)
}
```

# Control for batch effects

```{r}
# load table prepped from `bg_ballG_all_results.Rda` object
load("../processed/results/ballG_all_results/log_gfpkm_all.Rda")

# sample descriprions file
phenDat <- read.csv("../processed/describe_samples_batch.csv", stringsAsFactors = FALSE)
names(phenDat)[5]<-"batch"

# Prepare up data for sva
log_gfpkm_all <- as.matrix(log_gfpkm_all)
```

# sva model 1: without known batch

```{r}
# 1. sva
# Create a full model matrix without adjustment variables
  # since treatment has multiple levels, treat it as factor
mod = model.matrix(~as.factor(treatment) + as.factor(tissue), data=phenDat)

# Create null model (i.e. only adj. vars)
  # with no adjustment - just the intercept included
mod0 = model.matrix(~1, data=phenDat)

# Get number of batch variables
#n.sv = num.sv(basep_all, mod, method = "leek")
n.sv = num.sv(log_gfpkm_all, mod, method = "be", B = 5) # B are iterations
n.sv

# Note: B should be used only when method = "be"

# b) Estimage surrogate variables
svobj = sva(log_gfpkm_all, mod, mod0, n.sv=n.sv, B = 5)

plot(svobj$sv)
```

# sva model 2: known batch included only in full model

```{r}
mod2 <- model.matrix(~as.factor(treatment) + 
                              as.factor(tissue) + 
                              as.factor(batch), 
                              data=phenDat)

n.sv2 = num.sv(log_gfpkm_all, mod2, method = "be", B = 5)
#n.sv_adj

svobj2 = sva(log_gfpkm_all, mod2, mod0, n.sv=n.sv2, B = 5)

plot(svobj2$sv)
```

# sva model 3: include known batch in both full and null models
# ref: https://support.bioconductor.org/p/54719/

```{r}
mod03 = model.matrix(~as.factor(batch), data=phenDat) 
mod = mod2 
svobj3 = sva(log_gfpkm_all, mod, mod03) 

plot(svobj3$sv)
```




# svaseq
# svaseq: Moderated log link sva (i.e. log(g + c; c=1)

```{r}
# use untransformed data - svaseq takes care
load("../processed/results/ballG_all_results/nolog_gfpkm_all.Rda")
nolog_gfpkm_all <- gfpkm_all
```

```{r}
# svaseq model 4: without known batch (ref: Love et al 2016, bottom page 31-page 33)
mod4 = model.matrix(~as.factor(treatment) + 
                              as.factor(tissue),
                    data=phenDat)
n.sv4 = num.sv(nolog_gfpkm_all, mod4, method = "be", B = 5)

svaseq4 = svaseq(nolog_gfpkm_all, mod4, mod0,  
                 n.sv = n.sv4, controls = NULL, 
                 method = c("irw", "two-step","supervised"))
plot(svaseq4$sv)

# svaseq model 5: include known batch in both full and null models
mod = mod2 
mod05 = model.matrix(~as.factor(batch), data=phenDat) 
svobj5 = svaseq(nolog_gfpkm_all, mod, mod05) # compare with vfilter inluded
plot(svobj5$sv, pch=19,col="red")

# NOTE on sva output:
# pprob.gam: A vector of the posterior probabilities each gene is affected by heterogeneity
# pprob.b: A vector of the posterior probabilities each gene is affected by mod n.sv
```

# Did svaseq do better than without it?
```{r}
par(mfrow=c(2,1), mar=c(5,3,3,1))
stripchart(svobj5$sv[,1] ~ phenDat$batch, vertical=TRUE, main="SV1")
abline(h=0)

stripchart(svobj5$sv[,2] ~ phenDat$batch, vertical=TRUE, main="SV2")
abline(h=0)

stripchart(svobj5$sv[,3] ~ phenDat$batch, vertical=TRUE, main="SV3")
abline(h=0)

# plus 4 more
```

# Remove batch effects from sva results
# Note: this article cautions against adjusting pvals in svaseq

```{r}
# 1) Calculate F-test p-values for DE wrt diet, without adjusting for SVs
pValues_svaseq = f.pvalue(nolog_gfpkm_all, mod, mod0) #note mod used, not mod1
qValues_svaseq = p.adjust(pValues_svaseq, method = "BH") 

# 2) Adjust for surrogates
  # include surrogates in both full and null models, then get p-vals and q-vals
modSv_svaseq = cbind(mod, svobj5$sv) 
# sv is matrix whose columns are estimated SVs

mod0Sv_svaseq = cbind(mod03, svobj5$sv)

pValuesSv_svaseq = f.pvalue(nolog_gfpkm_all, modSv_svaseq, mod0Sv_svaseq)
qValuesSv_svaseq = p.adjust(pValuesSv_svaseq, method = "BH")

qres <- as.data.frame(qValuesSv_svaseq)
names(qres)[1]<-"qVal"
qresSig <- subset(qres, qVal < 0.1)
head(resSig[ order( resSig$log2FoldChange ), ])

# BUT, how to get adjusted expression values for downstream analysis?
```

# Path1: Add svs to sample variables

# https://support.bioconductor.org/p/87508/ ("If your downstream goal involves any type of significance test (differential analysis between groups for example) then we _do not_ recommend cleaning your data and then performing analysis on that cleaned data. This will cause problems with the calculation of the degrees of freeom for the statistical test. Instead we suggest incorporating the estimated surrogate variables as covariates in your analyses with limma, egdeR or DESeq after performing your svaseq analysis. J. Leek)

```{r}
phenDat$SV1 <- svobj5$sv[,1]
phenDat$SV2 <- svobj5$sv[,2]
phenDat$SV3 <- svobj5$sv[,3]
phenDat$SV4 <- svobj5$sv[,4]
phenDat$SV5 <- svobj5$sv[,5]
phenDat$SV6 <- svobj5$sv[,6]
phenDat$SV7 <- svobj5$sv[,7]

```

# Path2: Get corrected expression values

```{r}
# a) can use function
svaseq.dat <- cleaningY(nolog_gfpkm_all, mod2, svobj5)

# compare plots
plot(nolog_gfpkm_all[,c("C.1_H","C.2_H")])  # before sva
plot(svaseq.dat[,c("C.1_H","C.2_H")]) # after sva

# b) can use sva & limma
clean.dat <- svaseq(nolog_gfpkm_all, mod2, mod05)
Covar <- as.vector(cbind(clean.dat$sv[,1:8])) #if SVs >1; also try [,1:nsv]
#Covar <- as.factor(cbind(clean.dat$sv)) #if 1 SV
nolog_gfpkm_svaseq2 <- removeBatchEffect(clean.dat, covariates = Covar)
#plot(nolog_gfpkm_svaseq2)
```

# Expression pattern for batch-controlled data

#   svaseq.dat = expr values using f(x); 
#   svaseq_basep_all2 = expr values using limma

```{r}
#tfpkm_bg <-log2(svaseq.dat+1)
colnames(svaseq.dat) <- phenDat$id
tbg<- as.data.frame(svaseq.dat) %>% gather(id, fpkm)
tbg$treatment <- rep(c("C","DR","HS"), each=18*nrow(svaseq.dat))
tbg$tissue <- rep(c("B","H","O"),each=nrow(svaseq.dat),6)
tbg$replicate <- rep(c(1,2,3,4,5,6),each=nrow(svaseq.dat)*3)
tbg$prep <- paste(tbg$treatment,tbg$replicate,sep="_")
tbg$trep <- paste(tbg$tissue,tbg$replicate,sep="_")

b1 <- ggplot(tbg, aes(x=as.factor(prep), y=fpkm,color=treatment)) +
  geom_boxplot(alpha=0.9) +
          theme(axis.title.y = element_text(size = 24),
                axis.title.x=element_blank(),
                axis.text.x  = element_text(angle=90, vjust=0.5),
                text = element_text(size = 20))+
                
  scale_color_manual(values=c("#CC6666", "#9999CC", "#66CC99")) +
  ylab(expression("log"[2]*"(FPKM + 1)")) +
  xlab("Samples")+
  facet_grid(tissue~.)
b1 + theme(legend.position="none")
  
#ggsave(filename = "../plots/Expression_by_sample_batchControlled.pdf",width = 5,height=5)

b2 <- ggplot(tbg, aes(x=as.factor(trep), y=fpkm,color=tissue)) +
  geom_boxplot(alpha=0.9) +
          theme(axis.title.y = element_text(size = 24),
                axis.title.x=element_blank(),
                axis.text.x  = element_text(angle=90, vjust=0.5),
                text = element_text(size = 20)) +
  scale_color_manual(values=c("#CC6666", "#9999CC", "#66CC99")) +
  ylab(expression("log"[2]*"(FPKM + 1)")) +
  xlab("Tissue") +
  #scale_fill_discrete(name="Samples")+
  facet_grid(treatment~.)
b2 + theme(legend.position="none")
```

## PCA for global patterns - batch controlled data

```{r}
# all samples
#gfpkm_bg <-gexpr(bg_ballG_filt)
#gfpkm_bg <-log2(gfpkm_bg +1)
#colnames(gfpkm_bg)<-phenDat$id
#save(svaseq.dat, file = "../processed/results/ballG_all_results/svaseq_dat.Rda")

X<- t(scale(t(tbg),center=TRUE,scale=FALSE))
sv<- svd(t(X))
U<- sv$u
V<- sv$v
D<- sv$d
Z<- t(X)%*%V

# put PC1 and PC2 into a dataframe and add the phenotpe info side by side
pc.dat<- data.frame(phenDat$id, PC1 = Z[,1], PC2 = Z[,2], PC3 = Z[,3],PC4 =Z[,4])
s.id <- strsplit(as.character(pc.dat$phenDat.id), split="-",fixed=TRUE)
pc.dat$Diet <-unlist(lapply(s.id, function(x) x[1]))
pc.dat$Tissue <- unlist(lapply(strsplit(as.character(phenDat$id), 
                                        split="_", fixed=TRUE), function(x) x[2]))

# plot with ggplot2
a <- ggplot(pc.dat, aes(x=PC1, y=PC2, col=Diet, pch=Tissue)) + 
        geom_point(alpha=0.5, size=5) + 
        scale_color_manual(values=c("#CC6666", "#9999CC", "#66CC99")) +
        theme_bw() +
        theme(axis.title = element_text(size = 24)) +
        theme(text = element_text(size = 20)) +
        theme(panel.border = element_blank(), 
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), 
              axis.line.x = element_line(color="black", size = 0.8),
              axis.line.y = element_line(color="black", size = 0.8))
a 

ggsave(filename = "../plots/PCA_overall.pdf",width = 5,height=5)
```

