---
title: "Concensus coexpression network"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(79103)
```

# Categorical variables - Ref. point5 in:
# https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/faq.html

```{r}

library(WGCNA);
options(stringsAsFactors = FALSE);
enableWGCNAThreads()

#Read in the female liver data set
femData = read.csv("LiverFemale3600.csv");

# Read in the male liver data set
maleData = read.csv("LiverMale3600.csv");

# Take a quick look 
dim(femData)
names(femData)
dim(maleData)
names(maleData)
```

# Remove unwanted variables
# Put data in a multi-set format

```{r}
# We work with two sets:
nSets = 2;

# For easier labeling of plots, create a vector holding descriptive names of the two sets.
setLabels = c("Female liver", "Male liver")
shortLabels = c("Female", "Male")

# Form multi-set expression data: columns starting from 9 contain actual expression data.
multiExpr = vector(mode = "list", length = nSets)

multiExpr[[1]] = list(data = as.data.frame(t(femData[-c(1:8)])));
names(multiExpr[[1]]$data) = femData$substanceBXH;
rownames(multiExpr[[1]]$data) = names(femData)[-c(1:8)];

multiExpr[[2]] = list(data = as.data.frame(t(maleData[-c(1:8)])));
names(multiExpr[[2]]$data) = maleData$substanceBXH;
rownames(multiExpr[[2]]$data) = names(maleData)[-c(1:8)];

# Check that the data has the correct format for many functions operating on multiple sets:
exprSize = checkSets(multiExpr)
```

# Data cleaning & removal of outliers

```{r}
# Check that all genes and samples have sufficiently low numbers of missing values.
gsg = goodSamplesGenesMS(multiExpr, verbose = 3);
gsg$allOK
```

# If the last statement returns FALSE, remove offending samples & genes

```{r}
if (!gsg$allOK)
{
  # Print information about the removed genes:
  if (sum(!gsg$goodGenes) > 0)
    printFlush(paste("Removing genes:", paste(names(multiExpr[[1]]$data)[!gsg$goodGenes], 
                                              collapse = ", ")))
  for (set in 1:exprSize$nSets)
  {
    if (sum(!gsg$goodSamples[[set]]))
      printFlush(paste("In set", setLabels[set], "removing samples",
                       paste(rownames(multiExpr[[set]]$data)[!gsg$goodSamples[[set]]], collapse = ", ")))
    # Remove the offending genes and samples
    multiExpr[[set]]$data = multiExpr[[set]]$data[gsg$goodSamples[[set]], gsg$goodGenes];
  }
  # Update exprSize
  exprSize = checkSets(multiExpr)
}
```

# Cluster the samples on their Euclidean distance, separately in each set

```{r}
sampleTrees = list()

for (set in 1:nSets)
{
  sampleTrees[[set]] = hclust(dist(multiExpr[[set]]$data), method = "average")
}
```

# View sample trees together

```{r}
pdf(file = "SampleClustering.pdf", width = 12, height = 12);
par(mfrow=c(2,1))
par(mar = c(0, 4, 2, 0))

for (set in 1:nSets)
  plot(sampleTrees[[set]], main = paste("Sample clustering on all genes in", setLabels[set]),
       xlab="", sub="", cex = 0.7);
dev.off()
```

# Remove the 1 outlier in female data set

```{r include=FALSE}
# Choose the "base" cut height for the female data set
baseHeight = 16

# Adjust the cut height for the male data set for the number of samples
cutHeights = c(16, 16*exprSize$nSamples[2]/exprSize$nSamples[1]);

# Re-plot the dendrograms including the cut lines
pdf(file = "SampleClustering1.pdf", width = 12, height = 12);
par(mfrow=c(2,1))
par(mar = c(0, 4, 2, 0))
for (set in 1:nSets)
{
  plot(sampleTrees[[set]], main = paste("Sample clustering on all genes in", setLabels[set]),
       xlab="", sub="", cex = 0.7);
  abline(h=cutHeights[set], col = "red");
}
dev.off()

# Remove outlier

for (set in 1:nSets)
{
  # Find clusters cut by the line
  labels = cutreeStatic(sampleTrees[[set]], cutHeight = cutHeights[set])
  # Keep the largest one (labeled by the number 1)
  keep = (labels==1)
  multiExpr[[set]]$data = multiExpr[[set]]$data[keep, ]
}
collectGarbage();

# Check the size of the leftover data
exprSize = checkSets(multiExpr)
exprSize
```

# Clinical trait data

```{r}
traitData = read.csv("ClinicalTraits.csv");
dim(traitData)
names(traitData)

# remove columns that hold information we do not need.
allTraits = traitData[, -c(31, 16)];
allTraits = allTraits[, c(2, 11:36) ];

# See how big the traits are and what are the trait and sample names
dim(allTraits)
names(allTraits)
allTraits$Mice

# Form a multi-set structure that will hold the clinical traits.
Traits = vector(mode="list", length = nSets);
for (set in 1:nSets)
{
  setSamples = rownames(multiExpr[[set]]$data);
  traitRows = match(setSamples, allTraits$Mice);
  Traits[[set]] = list(data = allTraits[traitRows, -1]);
  rownames(Traits[[set]]$data) = allTraits[traitRows, 1];
}
collectGarbage();

# Define data set dimensions
nGenes = exprSize$nGenes;
nSamples = exprSize$nSamples;
```

```{r}
save(multiExpr, Traits, nGenes, nSamples, setLabels, shortLabels, exprSize, 
     file = "Consensus-dataInput.RData")

```


# One-step automatic network construction and module detection


```{r}
# Load the data saved in the first part
lnames = load(file = "Consensus-dataInput.RData")
lnames

# Get the number of sets in the multiExpr structure
nSets = checkSets(multiExpr)$nSets
```

# One-step network construction and module detection

```{r}
# Choose a set of soft-thresholding powers
powers = c(seq(4,10,by=1), seq(12,20, by=2));

# Initialize a list to hold the results of scale-free analysis
powerTables = vector(mode = "list", length = nSets);

# Call the network topology analysis function for each set in turn
for (set in 1:nSets)
  powerTables[[set]] = list(data = pickSoftThreshold(multiExpr[[set]]$data,
                                                     powerVector=powers,
                                                     verbose = 2)[[2]]);
collectGarbage();

# Plot the results:
colors = c("black", "red")

# Will plot these columns of the returned scale free analysis tables
plotCols = c(2,5,6,7)
colNames = c("Scale Free Topology Model Fit", "Mean connectivity", "Median connectivity",
             "Max connectivity");

# Get the minima and maxima of the plotted points
ylim = matrix(NA, nrow = 2, ncol = 4);
for (set in 1:nSets)
{
  for (col in 1:length(plotCols))
  {
    ylim[1, col] = min(ylim[1, col], powerTables[[set]]$data[, plotCols[col]], na.rm = TRUE);
    ylim[2, col] = max(ylim[2, col], powerTables[[set]]$data[, plotCols[col]], na.rm = TRUE);
  }
}

# Plot the quantities in the chosen columns vs. the soft thresholding power
sizeGrWindow(8, 6)
pdf(file = "scaleFreeAnalysis.pdf", wi = 8, he = 6)
par(mfcol = c(2,2));
par(mar = c(4.2, 4.2 , 2.2, 0.5))
cex1 = 0.7;
for (col in 1:length(plotCols)) for (set in 1:nSets)
{
  if (set==1)
  {
    plot(powerTables[[set]]$data[,1], -sign(powerTables[[set]]$data[,3])*powerTables[[set]]$data[,2],
         xlab="Soft Threshold (power)",ylab=colNames[col],type="n", ylim = ylim[, col],
         main = colNames[col]);
    addGrid();
  }
  if (col==1)
  {
    text(powerTables[[set]]$data[,1], -sign(powerTables[[set]]$data[,3])*powerTables[[set]]$data[,2],
         labels=powers,cex=cex1,col=colors[set]);
  } else
    text(powerTables[[set]]$data[,1], powerTables[[set]]$data[,plotCols[col]],
         labels=powers,cex=cex1,col=colors[set]);
  if (col==1)
  {
    legend("bottomright", legend = setLabels, col = colors, pch = 20) ;
  } else
    legend("topright", legend = setLabels, col = colors, pch = 20) ;
}
dev.off();
```

# Network construction and consensus module detection

```{r}
# Choose the power 6 for both sets
net = blockwiseConsensusModules(
  multiExpr, power = 6, minModuleSize = 30, deepSplit = 2,
  pamRespectsDendro = FALSE, 
  mergeCutHeight = 0.25, numericLabels = TRUE,
  minKMEtoStay = 0,
  saveTOMs = TRUE, verbose = 5)

```

# Look at detected modules

```{r}
names(net)

consMEs = net$multiMEs;
moduleLabels = net$colors;

# Convert the numeric labels to color labels
moduleColors = labels2colors(moduleLabels)
consTree = net$dendrograms[[1]] 
```

# Plot results

```{r}
sizeGrWindow(8,6);
pdf(file = "ConsensusDendrogram-auto.pdf", wi = 8, he = 6)
plotDendroAndColors(consTree, moduleColors,
                    "Module colors",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Consensus gene dendrogram and module colors")

dev.off()
```

```{r}
save(consMEs, moduleLabels, moduleColors, consTree, file = "Consensus-NetworkConstruction-auto.RData")

```


