---
title: "tutorial for removing batch effects in RNA data"
author: "Enoch Ng'oma"
date: "June 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sva)
library(bladderbatch)
data(bladderdata)
library(pamr)
library(limma)
```

# Set up data for sva

```{r}
# 1. Data should be feature-by-sample matrix
# 2. sva assumes presence of 2 variable types: bio-vars, 2) covariate vars
# 3. Needs 2 matrices: 
  # 1) full model - terms for adjustment variables and variables of interest; 
  # 2) null model - terms for all adjustment variables but not variables of interest
# i.e. analyze assoc. btn vars of interest and gene expr, adjusting for adjustment vars


# 1. create model matrices using model.matrix()

  # variable of interest: cancer status
  # data stored in ExpressionSet

# Obtain phenotype variables 
pheno = pData(bladderEset)

# Obtain expression variables
edata = exprs(bladderEset)

# Create a full model matrix
# at this point we have only var of interest, no adj. vars
# since cancer status has multiple levels, treat it as factor
mod = model.matrix(~as.factor(cancer), data=pheno)
```

# Apply sva() to estimate batch and other artifacts

```{r}
# Null model (i.e. only adj. vars)
# Since we are not adj. for any vars now, only an intercept is included
mod0 = model.matrix(~1, data=pheno)

# a) get number of latent factors
n.sv = num.sv(edata, mod, method = "leek")
n.sv

# b) estimage surrogate variables
svobj = sva(edata, mod, mod0, n.sv=n.sv)

# Note on sv object:
# The function returns a list with four components, sv, pprob.gam, pprob.b, n.sv. sv is a matrix whose columns correspond to the estimated surrogate variables. They can be used in downstream analyses as described below. pprob.gam is the posterior probability that each gene is associated with one or more latent variables [?]. pprob.b is the posterior probability that each gene is associated with the variables of interest [?]. n.sv is the number of surrogate variables estimated by the sva.
```

# Adjusting for surrogate variables - f.pvalue()

```{r}
# f.pvalue() calculates parametric F-test p-values for EACH row of data matrix
# F-test compares mod and mod0
# They MUST be nested models (i.e. all variables in mod0 must appear in mod)

# 1) First, calculate F-test p-values for differential expression with respect to cancer status, without adjusting for surrogate variables, adjust them for multiple testing, and calculate the number that are significant with a Q-value less than 0.05.
pValues = f.pvalue(edata, mod, mod0)
qValues = p.adjust(pValues, method = "BH") # BH, Benjamini-Hochberg for FDR

# Note that nearly 70% of the genes are strongly di erentially expressed at an FDR of less than 5% between groups. This number seems artificially high, even for a strong phenotype like cancer

# 2) Adjust for surrogates

# a) include surrogates in both full and null models, then get p-vals and q-vals
modSv = cbind(mod, svobj$sv) # sv is matrix whose columns are estimated SVs
mod0Sv = cbind(mod0, svobj$sv)
pValuesSv = f.pvalue(edata, modSv, mod0Sv)
qValuesSv = p.adjust(pValuesSv, method = "BH")

# These are the adjusted P-values and Q-values accounting for surrogate variables.
```

# Adjusting for surrogate variables using the limma package

```{r}
# 1) fit a linear model with surrogate variables
fit = lmFit(edata, modSv)

#From here, you can use the limma functions to perform the usual analyses.
```

# Suppose we want to calculate DE wrt cancer ( with limma package)


```{r}
# 1) compute contrasts between pairs of cancer/normal terms
    # no surrogates included here since SVs are only useful for adjustment
contrast.matrix <- cbind("C1"=c(-1,1,0,rep(0,svobj$n.sv)),
                         "C2"=c(0,-1,1,rep(0,svobj$n.sv)),
                         "C3"=c(-1,0,1,rep(0,svobj$n.sv)))

fitContrasts = contrasts.fit(fit,contrast.matrix) 

# 2) calculate the test statistic using eBayes()
eb = eBayes(fitContrasts)
topTable(eb, adjust="BH")

```

# Adjusting for known batches using ComBat
# ComBat() adjusts for known batches using an empirical Bayesian framework

```{r}
# There must be a known batch variable in the data
batch = pheno$batch

# 1) Create a model matrix for adj. vars + bio-vars
# Note that you do not include batch in creating this model matrix - it will be included later in the ComBat function. 
# In this case there are no other adj. vars so we simply fit an intercept term.
modcombat = model.matrix(~1, data=pheno) # i.e. mod0 above
```

# Important
# Note that adjustment variables will be treated as given to the ComBat function. This means if you are trying to adjust for a categorical variable with p di erent levels, you will need to give p-1 indicator variables for this covariate. We recommend using the     function to set these up. For continuous adjustment variables, just give a vector in the containing the covariate values in a single column of the model matrix.

```{r}
# 2) Adjust for batch
combat_edata = ComBat(dat=edata, batch=batch, mod=modcombat, par.prior=TRUE, prior.plots=TRUE)

# An expression matrix with same dims as input, but adj. for batch is created.

# 3) Perform significance testing on the batch-corrected matrix using the model matrix and null model matrix as described before:
pValuesComBat = f.pvalue(combat_edata,mod,mod0)
qValuesComBat = p.adjust(pValuesComBat,method="BH")

# NOTE:
# par.prior=FALSE gives the nonparametric empirical Bayesian adj. (takes longer) # prior.plots=FALSE gives prior plots with black as kernel estimate of the empirical batch efect density and red as the parametric estimate.

plot(combat_edata)
```

# If only mild effects of batch are expected

# NOTE: 
# Use the mean.only=TRUE option, to adjust only the mean of the batch effects across batches (default adjusts the mean and variance). The mean-only option is recommended if milder batch effects are expected (so no need to adjust the variance), or in cases where the variances are expected to be di erent across batches due to the biology.

```{r}
combat_edata1 = ComBat(dat=edata, batch=batch, mod=modcombat, mean.only=TRUE, par.prior=TRUE, prior.plots=TRUE)

```

# Create a reference batch effect

```{r}
# add the ref.batch parameter - see tutorial (p.9) for rationale & application
```

# Removing known batch effects with a linear model












