---
title: "Untitled"
author: "Enoch Ng'oma"
date: "6/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sva)
library(pamr)
library(limma)
library(dplyr)
```

# STEP 1: Data - all transcripts across tissues and diets
# Genes in rows and samples/variables in columns
# Phenotype data as a separate file

```{r}
# read extracted table
basep_all <- read.csv("../processed/results/ballG_all_results/gene_associated_texpr.csv",
                    stringsAsFactors = FALSE)

phenDat <- read.csv("../processed/describe_samples_batch.csv", stringsAsFactors = FALSE)
```

# STEP 2: Set up data for sva
# sva assumes 2 variable types present: 
# 1) adjustment variables - groups (i.e. 6)
# 2) vars of interest - treatment and tissue

```{r}
# Create model matrices using model.matrix()

# Phenotype variables 
phenDat = phenDat 

# Only expression variables needed and should be in matrix form
basep_all = basep_all %>%
  select(-geneNames, -geneIDs) %>%
  select(tname, everything())

nams <- basep_all$tname
rownames(basep_all) = make.names(nams, unique=TRUE) 
basep_all$tname <- NULL

basep_all <- as.matrix(basep_all)

# Create a full model matrix without adjustment variables
  # since treatment has multiple levels, treat it as factor
mod = model.matrix(~as.factor(treatment) + as.factor(tissue), data=phenDat)

# Create null model (i.e. only adj. vars)
  # with no adjustment - just the intercept included (unsupervised)
mod0 = model.matrix(~1, data=phenDat)
```

# STEP 3: Apply unsupervised sva() to estimate batch and other artifacts

```{r}
# a) Get number of latent factors
#n.sv = num.sv(basep_all, mod, method = "leek")
n.sv = num.sv(basep_all, mod, method = "be", B = 5) # B are iterations
n.sv

# b) Estimage surrogate variables
svobj = sva(basep_all, mod, mod0, n.sv=n.sv, B = 5)

#svobj = sva(basep_all, mod, mod0) 
# errors, see:
# https://www.biostars.org/p/198820/
# What does this error mean? - issue implementing SVA
# https://stackoverflow.com/questions/43101585/error-when-generating-the-sva-object-using-package-sva-in-r
```

# STEP 4: Adjusting for surrogate variables - f.pvalue()

```{r}
# 1) First, calculate F-test p-values for DE wrt diet, without adjusting for SVs
pValues = f.pvalue(basep_all, mod, mod0)
qValues = p.adjust(pValues, method = "BH") 

# 2) Second, adjust for surrogates
  # include surrogates in both full and null models, then get p-vals and q-vals
modSv = cbind(mod, svobj$sv) # sv is matrix whose columns are estimated SVs
mod0Sv = cbind(mod0, svobj$sv)

pValuesSv = f.pvalue(basep_all, modSv, mod0Sv)
qValuesSv = p.adjust(pValuesSv, method = "BH")

# These are the adjusted P-values and Q-values accounting for surrogate variables.
```

# OPTIONAL STEP 4b: Adjusting for surrogate variables using the limma package

```{r}
# 1) fit a linear model with surrogate variables
fit = lmFit(basep_all, modSv)

# Can now use limma functions to perform the usual analyses.
# Suppose we want to calculate DE wrt diet:

# 1) compute contrasts between pairs of diet terms
    # no surrogates included here since SVs are only useful for adjustment
contrast.matrix <- cbind("C1"=c(-1,1,0,rep(0,svobj$n.sv)),
                         "C2"=c(0,-1,1,rep(0,svobj$n.sv)),
                         "C3"=c(-1,0,1,rep(0,svobj$n.sv)))

fitContrasts = contrasts.fit(fit,contrast.matrix) 

# 2) calculate the test statistic using eBayes()
eb = eBayes(fitContrasts)
topTable(eb, adjust="BH")
```

# ALTERNATIVE APPROACH: ADJUSTING FOR KNOWN BATCHES

# STEP 5: Apply the ComBat function
# ComBat() adjusts for known batches using an empirical Bayesian framework

```{r}
# 'group' is a known batch variable in the data
batch = phenDat$group

# 1) Create a model matrix for adj. vars + bio-vars
# Note that you do not include batch in creating this model matrix - it will be included later in the ComBat function. 
# In this case there are no other adj. vars so we simply fit an intercept term.
modcombat = model.matrix(~1, data=phenDat) # i.e. same as mod0 above

# 2) Adjust for batch
combat_basep_all = ComBat(dat=basep_all, 
                          batch=batch, 
                          mod=modcombat, 
                          par.prior=TRUE, 
                          prior.plots=TRUE)
# Output is an expression matrix with same dims as input, but adjusted for batch.

# 3) Perform significance testing on the batch-corrected matrix as before:
pValuesComBat = f.pvalue(combat_basep_all,mod,mod0)
qValuesComBat = p.adjust(pValuesComBat,method="BH")

# NOTE:
# par.prior=FALSE gives the nonparametric empirical Bayesian adj. (takes longer) 
# prior.plots=TRUE gives prior plots with black as kernel estimate of the empirical batch efect density and red as the parametric estimate.

plot(combat_basep_all)
```

# STEP 6 (optional): If only mild effects of batch are expected
# If variances expected to differ across batches due to biology

```{r}
# option mean.only=TRUE adjusts the mean of batch effects, not variance
combat_basep_all_mean = ComBat(dat=basep_all, 
                           batch=batch, 
                           mod=modcombat, 
                           mean.only=TRUE, 
                           par.prior=TRUE, 
                           prior.plots=TRUE)

```

# svaseq: Moderated log link sva
# modeling with variables of interest, known batch variables and unknown artifacts
 
```{r}
# unsupervised form
svaseq_basep_all <- svaseq(basep_all, mod, mod0 , n.sv, controls = NULL,
                           method = c("irw", "two-step", "supervised"), B = 5,
                           numSVmethod = "be")
svaseq_basep_all$sv

# supervised form
group = phenDat$group
mod1 = model.matrix(~as.factor(treatment) + as.factor(tissue) + as.factor(group), data=phenDat)

mod0 = cbind(mod1[,1])
svaseq_basep_all1 = svaseq(basep_all,mod1,mod0,n.sv=11)$sv

svaseq_basep_all1$sv

plot(svseq_base_all1, pch=19,col="red")
```

```{r}
par(mfrow=c(2,1),mar=c(3,5,3,1))
stripchart(svaseq_basep_all$sv[,c(1:10)] ~ phenDat$group,vertical=TRUE,main="SV")
abline(h=0)

stripchart(svaseq_basep_all$sv[,1] ~ phenDat$group,vertical=TRUE,main="SV1")
abline(h=0)

stripchart(svaseq_basep_all$sv[,2] ~ phenDat$group,vertical=TRUE,main="SV2")
abline(h=0)

stripchart(svaseq_basep_all$sv[,10] ~ phenDat$group,vertical=TRUE,main="SV10")
abline(h=0)

stripchart(svseq$sv[,2] ~ dds$cell,vertical=TRUE,main="SV2") abline(h=0)
```


# Estimate batch with pca

```{r}
#pca_basep_all = log(tfpkm + 1)
batch_pca = svd(basep_all - rowMeans(basep_all))$v[,1]
```


