---
title: "DiffExpr"
author: "Enoch Ng'oma"
date: "5/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#R
#load R version R version 3.2.2
#load R packagees:
library(ballgown)
library(genefilter)
library(tidyverse)
library(devtools)
library(cowplot)
library(reshape)
#library(ggthemes)

```

# Differential expression analysis protocol

```{r}
phenDat <- read.csv("describe_samples.csv", stringsAsFactors = FALSE)

#bg_ballG <- ballgown(dataDir = "../processed/ballG_all", samplePattern = "", pData = phenDat)

# Save .Rda
#save(bg_ballG, file = "../processed/results/ballG_all_results/bg_ballG_all_results.Rda")

load("../processed/results/ballG_all_results/bg_ballG_all_results.Rda")

# Dispplay a decsription
bg_ballG

```

# Filter to remove low-abundance genes (i.e. those with variance across samples of <=1)

```{r}
bg_ballG_filt <- subset(bg_ballG, "rowVars(texpr(bg_ballG)) >1", genomesubset=TRUE)

# See counts
bg_ballG_filt

save(bg_ballG_filt, file = "../processed/results/ballG_all_results/bg_ballG_filt.Rda")
```

## Overall expression numbers
# Distribution of gene abundances (FPKM values) across samples (libraries), colored by treatment or tissue and faceted 

```{r}
load(file = "../processed/results/ballG_all_results/bg_ballG_filt.Rda")

tfpkm_bg<-texpr(bg_ballG_filt, meas="FPKM")
tfpkm_bg <-log2(tfpkm_bg+1)
colnames(tfpkm_bg)<-phenDat$id
save(tfpkm_bg, file = "../processed/results/ballG_all_results/tfpkm_bg_ballG_filt.Rda")


tbg<- as.data.frame(tfpkm_bg) %>% gather(id, fpkm)
tbg$treatment <- rep(c("C","DR","HS"), each=18*nrow(tfpkm_bg))
tbg$tissue <- rep(c("B","H","O"),each=nrow(tfpkm_bg),6)
tbg$replicate <- rep(c(1,2,3,4,5,6),each=nrow(tfpkm_bg)*3)
tbg$prep <- paste(tbg$treatment,tbg$replicate,sep="_")
tbg$trep <- paste(tbg$tissue,tbg$replicate,sep="_")

b1 <- ggplot(tbg, aes(x=as.factor(prep), y=fpkm,fill=treatment)) +
  geom_boxplot() +
  ylab(expression("log"[2]*"(FPKM + 1)")) +
  facet_grid(tissue~.)
b1

b2 <- ggplot(tbg, aes(x=as.factor(trep), y=fpkm,fill=tissue)) +
  geom_boxplot() +
  ylab(expression("log"[2]*"(FPKM + 1)")) +
  facet_grid(treatment~.)
b2

```

## PCA for global patterns


```{r}

gfpkm_bg <-gexpr(bg_ballG_filt)
gfpkm_bg <-log2(gfpkm_bg +1)
colnames(gfpkm_bg)<-phenDat$id
save(gfpkm_bg, file = "../processed/results/ballG_all_results/gfpkm_bg_ballG_filt.Rda")


# all samples
X<- t(scale(t(gfpkm_bg),center=TRUE,scale=FALSE))
sv<- svd(t(X))
U<- sv$u
V<- sv$v
D<- sv$d
Z<- t(X)%*%V

# put PC1 and PC2 into a dataframe and add the phenotpe info side by side
pc.dat<- data.frame(phenDat$id, PC1 = Z[,1], PC2 = Z[,2], PC3 = Z[,3],PC4 =Z[,4])
s.id <- strsplit(as.character(pc.dat$phenDat.id), split="-",fixed=TRUE)
pc.dat$Diet <-unlist(lapply(s.id, function(x) x[1]))
pc.dat$Tissue <- unlist(lapply(strsplit(as.character(phenDat$id), split="_", fixed=TRUE), function(x) x[2]))
## make figure with ggplot2

#e <- rep(c('B','H','O'),each=18)
#name <- phenDat$tissue
a <- ggplot(pc.dat, aes(x=PC1, y=PC2, col=Diet, pch=Tissue)) + 
        geom_point(alpha=0.7, size=3) + 
        theme_bw() +
        theme(panel.border = element_blank(), 
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), 
              axis.line.x = element_line(color="black", size = 0.6),
              axis.line.y = element_line(color="black", size = 0.6))
a 

ggsave(filename = "../plots/PCA_overall.pdf",width = 5,height=5)
```

## Differential expression visualization

```{r}
load("/home/engoma/MyGithub/BasePop_RNAseq/base_pop/processed/results/bg_results_tissues/heads.Rda")
bg_ballG_heads_filt <- subset(bg_ballG_heads, "rowVars(texpr(bg_ballG_heads)) >1", genomesubset=TRUE)

load("/home/engoma/MyGithub/BasePop_RNAseq/base_pop/processed/results/bg_results_tissues/soma.Rda")
bg_ballG_ovaries_filt <- subset(bg_ballG_ovaries, "rowVars(texpr(bg_ballG_ovaries)) >1", genomesubset=TRUE)
load("/home/engoma/MyGithub/BasePop_RNAseq/base_pop/processed/results/bg_results_tissues/ovaries.Rda")
bg_ballG_soma_filt <- subset(bg_ballG_soma, "rowVars(texpr(bg_ballG_soma)) >1", genomesubset=TRUE)

heads_phen <- subset(phenDat, phenDat$tissue=="H")
soma_phen <- subset(phenDat, phenDat$tissue=="B")
ovaries_phen <- subset(phenDat, phenDat$tissue=="O")

h_gfpkm_bg <-gexpr(bg_ballG_heads_filt)
h_gfpkm_bg <-log2(h_gfpkm_bg +1)
colnames(h_gfpkm_bg)<-heads_phen$id

o_gfpkm_bg <-gexpr(bg_ballG_ovaries_filt)
o_gfpkm_bg <-log2(o_gfpkm_bg +1)
colnames(o_gfpkm_bg)<-ovaries_phen$id

s_gfpkm_bg <-gexpr(bg_ballG_soma_filt)
s_gfpkm_bg <-log2(s_gfpkm_bg +1)
colnames(s_gfpkm_bg)<-soma_phen$id

#read in diff exp objects
hd.r <- read.csv(file="../processed/results/bg_results_tissues/results_bg_heads_gene.csv", header=TRUE)

ov.r <- read.csv(file="../processed/results/bg_results_tissues/results_bg_ovaries_gene.csv", header=TRUE)

so.r <- read.csv(file="../processed/results/bg_results_tissues/results_bg_soma_gene.csv", header=TRUE)


rx <- as.character(hd.r[which(hd.r$qval<=0.05), 'id'])
h_ord <- h_gfpkm_bg[rx,]
res.sub <- hd.r[which(hd.r$qval<=0.05),]
t1 <- rowMeans(h_ord[rx,1:6])
t2<-rowMeans(h_ord[rx,7:12])
t3<-rowMeans(h_ord[rx,13:18])
tt <- abs(t1-t2)
v.set <-rowVars(t(scale(t(cbind(t1,t2,t3)),center=TRUE,scale=FALSE)))
m.set <- rowMeans(cbind(t1,t2,t3))
#& m.set <7 & m.set >3
ss <- which(res.sub$qval<0.009)
h_ord<-h_ord[ss,]
h_ord <- h_ord[order(rowMeans(h_ord[,1:6])),]

#h_ord <- h_ord[order(res.sub$qval),]
#h_ord <- h_ord[1:20,]

h_ord <- as.data.frame(h_ord)
h_ord$gname<-seq(1,nrow(h_ord))

hds.long<-melt(h_ord,id='gname')
#hds.long$variable <- unlist(lapply(strsplit(as.character(hds.long$variable), split="-",fixed=TRUE),function(x) x[1]))

ggplot(hds.long, aes(variable, gname)) +
  geom_tile(aes(fill = value), colour = "white") + 
  scale_fill_gradient(low = "white", high = "steelblue") +
  theme(#axis.text.x = element_blank(),
        axis.text.y=element_blank(),
        axis.title.x = element_blank(),
        axis.text.x  = element_text(angle=90, vjust=0.5),
        axis.title.y = element_blank())

######################
#SOMA

rx <- as.character(so.r[which(so.r$qval<=0.05), 'id'])
s_ord <- s_gfpkm_bg[rx,]
res.sub <- so.r[which(so.r$qval<=0.05),]
t1 <- rowMeans(s_ord[rx,1:6])
t2<-rowMeans(s_ord[rx,7:12])
t3<-rowMeans(s_ord[rx,13:18])
tt <- abs(t1-t2)
v.set <-rowVars(t(scale(t(cbind(t1,t2,t3)),center=TRUE,scale=FALSE)))
m.set <- rowMeans(cbind(t1,t2,t3))
#& m.set <7 & m.set >3
ss <- which(res.sub$qval<0.009 & m.set <7 & m.set >3)
s_ord<-s_ord[ss,]
s_ord <- s_ord[order(rowMeans(s_ord[,1:6])),]

#s_ord <- s_ord[order(res.sub$qval),]
#s_ord <- s_ord[1:20,]

s_ord <- as.data.frame(s_ord)
s_ord$gname<-seq(1,nrow(s_ord))

so.long<-melt(s_ord,id='gname')
#so.long$variable <- unlist(lapply(strsplit(as.character(so.long$variable), split="-",fixed=TRUE),function(x) x[1]))

ggplot(so.long, aes(variable, gname)) +
  geom_tile(aes(fill = value), colour = "white") + 
  scale_fill_gradient(low = "white", high = "steelblue") +
  theme(#axis.text.x = element_blank(),
        axis.text.y=element_blank(),
        axis.title.x = element_blank(),
        axis.text.x  = element_text(angle=90, vjust=0.5),
        axis.title.y = element_blank())

######################
#OVARY


rx <- as.character(ov.r[which(ov.r$qval<=0.05), 'id'])
o_ord <- o_gfpkm_bg[rx,]
res.sub <- ov.r[which(ov.r$qval<=0.05),]
t1 <- rowMeans(o_ord[rx,1:6])
t2<-rowMeans(o_ord[rx,7:12])
t3<-rowMeans(o_ord[rx,13:18])
tt <- abs(t1-t2)
v.set <-rowVars(t(scale(t(cbind(t1,t2,t3)),center=TRUE,scale=FALSE)))
m.set <- rowMeans(cbind(t1,t2,t3))

#ss <- which(res.sub$qval<0.009 & m.set <7 & m.set >3)
#o_ord<-o_ord[ss,]
o_ord <- o_ord[order(rowMeans(o_ord[,1:6])),]
o_ord<-o_ord[1:(nrow(o_ord)-1),]
#o_ord <- o_ord[order(res.sub$qval),]
#o_ord <- o_ord[1:20,]

o_ord <- as.data.frame(o_ord)
o_ord$gname<-seq(1,nrow(o_ord))

ov.long<-melt(o_ord,id='gname')
#ov.long$variable <- unlist(lapply(strsplit(as.character(ov.long$variable), split="-",fixed=TRUE),function(x) x[1]))

ggplot(ov.long, aes(variable, gname)) +
  geom_tile(aes(fill = value), colour = "white") + 
  scale_fill_gradient(low = "white", high = "steelblue") +
  theme(#axis.text.x = element_blank(),
        axis.text.y=element_blank(),
        axis.title.x = element_blank(),
        axis.text.x  = element_text(angle=90, vjust=0.5),
        axis.title.y = element_blank())

```



